```{r date_time}
today <- Sys.Date()
class(today)
now <- Sys.time()
class(now)
my_date <- as.Date("1971-05-14")

my_date + 1 # "1971-05-15"

my_date <- as.Date("1971-14-05", format = "%Y-%d-%m")

# Utc
my_dateposix <- as.POSIXct("1971-05-14 11:25:15")

# Date arithmetic

my_dateposix + 1 # "1971-05-14 11:25:16"

my_date2 <- as.Date("1998-09-29")

my_date2 - my_date # Time difference of 10000 days

# POSIXct arithmetic
my_time <- as.POSIXct("1971-05-14 11:25:15 UTC")

my_time + 1 # "1971-05-14 11:25:16 UTC"

my_time2 <- as.POSIXct("1974-07-14 21:11;55 UTC")
my_time2 - my_time # Time difference of 1157.407 days

# Under the hood it counts the number of days since 1970-01-01
my_date <- "1971-05-14"
unclass(my_date) # 498

# POSIX number of seconds since 1970-01-01 at midnight
my_time <- "1971-05-14 11:25:15 CET"
unclass(my_time) #43064715
attr(my_time, "tzone")
```
# Date/Time R Packages
- lubridate
- zoo
- xts

Create and format dates
To create a Date object from a simple character string in R, you can use the as.Date() function. The character string has to obey a format that can be defined using a set of symbols (the examples correspond to 13 January, 1982):

%Y: 4-digit year (1982)
%y: 2-digit year (82)
%m: 2-digit month (01)
%d: 2-digit day of the month (13)
%A: weekday (Wednesday)
%a: abbreviated weekday (Wed)
%B: month (January)
%b: abbreviated month (Jan)
The following R commands will all create the same Date object for the 13th day in January of 1982:

as.Date("1982-01-13")
as.Date("Jan-13-82", format = "%b-%d-%y")
as.Date("13 January, 1982", format = "%d %B, %Y")
Notice that the first line here did not need a format argument, because by default R matches your character string to the formats "%Y-%m-%d" or "%Y/%m/%d".

In addition to creating dates, you can also convert dates to character strings that use a different date notation. For this, you use the format() function. Try the following lines of code:

today <- Sys.Date()
format(Sys.Date(), format = "%d %B, %Y")
format(Sys.Date(), format = "Today is a %A!")

```{r excercise}
# Get the current date: today
today <- Sys.Date()

# See what today looks like under the hood
unclass(today)

# Get the current time: now
now <- Sys.time()

# See what now looks like under the hood
unclass(now)

# Definition of character strings representing times
str1 <- "May 23, '96 hours:23 minutes:01 seconds:45"
str2 <- "2012-3-12 14:23:08"

# Convert the strings to POSIXct objects: time1, time2
time1 <- as.POSIXct(str1, format = "%B %d, '%y hours:%H minutes:%M seconds:%S")
time2 <- as.POSIXct(str2, format = "%Y-%m-%d %H:%M:%S")

# Convert times to formatted strings
format(time1, "%M")
format(time2, "%I:%M %p")
```
```{r create_format_dates_1}
# Definition of character strings representing dates
str1 <- "May 23, '96"
str2 <- "2012-03-15"
str3 <- "30/January/2006"

# Convert the strings to dates: date1, date2, date3
date1 <- as.Date(str1, format = "%b %d, '%y")
date2 <- as.Date(str2)
date3 <- as.Date(str3, format = "%d/%B/%Y")


# Convert dates to formatted strings
format(date1, "%A")
format(date2, "%d")
format(date3, "%b %Y")
```

Create and format times
Similar to working with dates, you can use as.POSIXct() to convert from a character string to a POSIXct object, and format() to convert from a POSIXct object to a character string. Again, you have a wide variety of symbols:

%H: hours as a decimal number (00-23)
%I: hours as a decimal number (01-12)
%M: minutes as a decimal number
%S: seconds as a decimal number
%T: shorthand notation for the typical format %H:%M:%S
%p: AM/PM indicator
For a full list of conversion symbols, consult the strptime documentation in the console:

?strptime
Again,as.POSIXct() uses a default format to match character strings. In this case, it's %Y-%m-%d %H:%M:%S. In this exercise, abstraction is made of different time zones.

```{r create_formate_times_1}
# Definition of character strings representing times
str1 <- "May 23, '96 hours:23 minutes:01 seconds:45"
str2 <- "2012-3-12 14:23:08"

# Convert the strings to POSIXct objects: time1, time2
time1 <- as.POSIXct(str1, format = "%B %d, '%y hours:%H minutes:%M seconds:%S")
time2 <- as.POSIXct(str2, format = "%Y-%m-%d %H:%M:%S")

# Convert times to formatted strings
format(time1, "%M")
format(time2, "%I:%M %p")
```
?diff # diff function to calculate difference between consecutive elements in the vector

```
```{r dates_arithmetic}
# www.timeanddate.com
seasons <- c("spring", "summer", "fall", "winter")
astro <- c("20-Mar-2015", "25-Jun-2015", "23-Sep-2015", "22-Dec-2015")
meteo <- c("March 1, 15", "June 1, 15", "September 1, 15", "December 1, 15")
names(astro) <- seasons
names(meteo) <- seasons

# Convert astro to vector of Date objects: astro_dates
astro_dates <- as.Date(astro, format = "%d-%b-%Y")

# Convert meteo to vector of Date objects: meteo_dates
meteo_dates <- as.Date(meteo, format = "%B %d, %y")

# Calculate the maximum absolute difference between astro_dates and meteo_dates
max(abs(astro_dates - meteo_dates))
```